---
title: "An Experiment on Projection Bias in Effort Choices"
author: |
  | Marc Kaufmann (Central European University)
date: "December 2021"
output:
  beamer_presentation:
    theme: metropolis
header-includes: 
  - \usepackage{graphicx}
  - \usepackage{tcolorbox}
  - \usepackage{import}
fontsize: 10pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

**Research Question**: Do people exhibit projection bias in effort choices?

**Today**: How to create differences in marginal disutility of effort

Idea of the pilot that I ran in September 2021:

1. Everyone does required tasks
2. Two treatments:
    - *No Rest*: elicit WTW for 5, 8, 11, 15 additional tasks, then rest
    - *Rest*: rest, then elicit WTW for 5, 8, 11, 15 additional tasks
    
## Introduction

Projection bias predicts that people mispredict their future tastes as closer to their current tastes than they are.

- In effort choices: project their current taste for work ($D'(s)$) onto the future: 
\begin{equation*}
  \tilde{D}'(e|s) = (1 - \alpha) D'(e) + \alpha D'(s)
\end{equation*}

where $\alpha \in [0, 1]$ is degree of projection bias. 

For identification, need change in $D'(s)$, the current (marginal) disutility of work.

## Literature

Bla.

## Experiment Pilot: Design

Participants do the following:

- Tutorial with 2 tasks
- Some questions and consent
- After consent, get randomized into individually balanced (not cross-balanced) treatments:
    - *rest* or *no rest* before elicitation of WTW
    - 5, 10, or 15 required tasks
    - participation fee (can be high or low)
    - *classical music*, *meditation*, or *wave sounds* during rest (always 5-7 minutes)
- Required tasks
- *No Rest*: price list for WTW for 5, 8, 11, 15 extra tasks; then rest
- *Rest*: rest; then price list for WTW 
- Debrief survey

## Experiment Pilot: Summary Statistics

```{r load_data, message=FALSE, warning=FALSE, echo=FALSE}
library(tidyverse)
library(ggplot2)
library(kableExtra)

df <- readRDS("clean_df.rds")
```

```{r summary_stats, message=FALSE, echo=FALSE}
df %>%
  group_by(rest_treatment, relax_treatment) %>%
  summarise(n()) %>%
  kable()
```

## Experiment Pilot: Summary Statistics

```{r summary_stats2, echo=FALSE, message=FALSE}
df %>%
  group_by(rest_treatment, participation_fee) %>%
  summarise(n()) %>%
  kable()
```

## Experiment Pilot: Summary Statistics

```{r balanced3, echo=FALSE, message=FALSE}
df %>%
  group_by(rest_treatment, relax_treatment) %>%
  summarise(n()) %>%
  kable()
```

## Experiment Pilot: Results

```{r line_plot_main_treatment, echo=FALSE, message=FALSE}
df %>% 
  group_by(rest_treatment, extra_tasks) %>%
  summarize(WTW = mean(WTW), N = n()) %>%
  ggplot(mapping = aes(x = extra_tasks, y = WTW)) + 
  geom_line(aes(group = rest_treatment, color = rest_treatment)) + xlim(5, 15) + ylim(0, 3.2)
```

## Experiment Pilot: Results

```{r line_plot}
df %>%
  mutate(rest_treatment = if_else(rest_treatment == "elicit-then-get-rest", "NR", "R")) %>%
  mutate(cross = paste(rest_treatment, relax_treatment, sep = "_")) %>%
  group_by(cross, relax_treatment, rest_treatment, extra_tasks) %>%
  summarize(WTW = mean(WTW), N = n()) %>%
  ggplot(mapping = aes(x = extra_tasks, y = WTW, group = cross)) +
  geom_line(aes(color = relax_treatment, linetype = rest_treatment)) + xlim(5, 15) + ylim(0, 3.2) + 
  labs(title = "Splitting by rest treatment and type")
```

## Experiment Pilot: Results

```{r box_plot_main_treatments}
df %>%
  mutate(extra_tasks = as_factor(extra_tasks)) %>%
  ggplot(aes(x = extra_tasks, y = WTW)) + 
  geom_boxplot(aes(color = rest_treatment)) +
  facet_grid(. ~ rest_treatment)
```

## Experiment Pilot: Results

```{r box_plot_by_rest_type_no_rest}
df %>%
  filter(rest_treatment == "elicit-then-get-rest") %>%
  mutate(extra_tasks = as_factor(extra_tasks)) %>%
  ggplot(aes(x = extra_tasks, y = WTW)) + 
  geom_boxplot(aes(color = relax_treatment)) +
  facet_grid(. ~ relax_treatment) +
  labs(title = "No Rest before Elicitation")
```

## Experiment Pilot: Results

```{r box_plot_by_rest_type_rest}
df %>%
  filter(rest_treatment == "get-rest-then-elicit") %>%
  mutate(extra_tasks = as_factor(extra_tasks)) %>%
  ggplot(aes(x = extra_tasks, y = WTW)) + 
  geom_boxplot(aes(color = relax_treatment)) +
  facet_grid(. ~ relax_treatment) +
  labs(title = "Rest before Elicitation")
```

## Experiment Pilot: Results

\import{./}{basic_regs.tex}

## Experiment Pilot: Results

\import{./}{required_tasks_regs.tex}

## Experiment Pilot: Results

\import{./}{fee_regs.tex}

## Conclusion

How to get people tired -- how to get changes in marginal disutility?

- Across days?
- Within days?
- Without attrition (earlier experiment had that issue -- and no meaningful change in $D'$)

## Every End is a Beginning

\pause{}

\begin{center}
    \huge
    Thanks
    \normalsize
\end{center}

Email questions, compliments, and helpful comments to Marc: KaufmannM@ceu.edu


# References

## References
